Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{London version \YellowPaperVersionNumber}}

### 2023-11-27 Paper.tex
add floor to gas target calculation
adjust gas limit verification at the London fork block

\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
\begin{eqnarray}
& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
\nonumber& & H_{\mathrm{l}} \geqslant 5000
\end{eqnarray}

@@

where:
  \begin{align}
    \tau &\equiv \frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{\rho} \\
    \rho &\equiv 2 \\
    \nu^* &\equiv \begin{cases}
      \lfloor \frac{{P(H)_{\mathrm{H}}}_{\mathrm{f}} \times ({\tau - P(H)_{\mathrm{H}}}_{\mathrm{g}})}{\tau} \rfloor &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} < \tau \\
      \lfloor \frac{{P(H)_{\mathrm{H}}}_{\mathrm{f}} \times ({P(H)_{\mathrm{H}}}_{\mathrm{g}} - \tau)}{\tau} \rfloor &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau \\
  \end{cases} \\
  \nu &\equiv \begin{cases}
    \lfloor \frac{\nu^*}{\xi} \rfloor &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} < \tau \\
    max(\lfloor \frac{\nu^*}{\xi} \rfloor, 1) &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau \\
  \end{cases} \\
  \xi &\equiv 8
\end{align}

@@

@@

\hypertarget{block_header_validity_function}{}Thus we are able to define the block header validity function $V(H)$:
\begin{eqnarray}
V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
\nonumber & & H_{\mathrm{d}} = D(H) \quad \wedge \\
\nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
\nonumber& & H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}} \quad \wedge \\
\nonumber& & H_{\mathrm{i}} = {P(H)_{\mathrm{H}}}_{\mathrm{i}} +1 \quad \wedge \\
\nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32 \quad \wedge \\
\nonumber& & H_{\mathrm{f}} = F(H)
\end{eqnarray}
where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$

>>>>

@@

where:
  \begin{align}
    \tau &\equiv \lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{\rho} \rfloor \\
    \rho &\equiv 2 \\
    \nu^* &\equiv \begin{cases}
      \lfloor \frac{{P(H)_{\mathrm{H}}}_{\mathrm{f}} \times ({\tau - P(H)_{\mathrm{H}}}_{\mathrm{g}})}{\tau} \rfloor &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} < \tau \\
      \lfloor \frac{{P(H)_{\mathrm{H}}}_{\mathrm{f}} \times ({P(H)_{\mathrm{H}}}_{\mathrm{g}} - \tau)}{\tau} \rfloor &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau \\
  \end{cases} \\
  \nu &\equiv \begin{cases}
    \lfloor \frac{\nu^*}{\xi} \rfloor &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} < \tau \\
    max(\lfloor \frac{\nu^*}{\xi} \rfloor, 1) &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau \\
  \end{cases} \\
  \xi &\equiv 8
\end{align}

@@

\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
\begin{eqnarray}
& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}'} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor \quad \wedge \\
\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}'} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor \quad \wedge \\
\nonumber& & H_{\mathrm{l}} \geqslant 5000
\end{eqnarray}

where:
  \begin{align}
    {P(H)_{\mathrm{H}}}_{\mathrm{l}'} &\equiv \begin{cases}
      {P(H)_{\mathrm{H}}}_{\mathrm{l}} \times \rho &  \text{if} \quad H_{\mathrm{i}} = F_{\mathrm{London}} \\
      {P(H)_{\mathrm{H}}}_{\mathrm{l}} &  \text{if} \quad H_{\mathrm{i}} > F_{\mathrm{London}} \\
  \end{cases}
\end{align}

To avoid a discontinuity in gas usage, the value of the parent block gas limit for the purpose of validating the current block's gas limit is modified at the \textit{London} fork block by multiplying it by the \textit{elasticity multiplier}, $\rho$. We call this modified value ${P(H)_{\mathrm{H}}}_{\mathrm{l}'}$. This ensures that the gas target for post-\textit{London} blocks can be set roughly in line with the gas limit of pre-\textit{London} blocks.

@@

\hypertarget{block_header_validity_function}{}Thus we are able to define the block header validity function $V(H)$:
\begin{eqnarray}
V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
\nonumber & & H_{\mathrm{d}} = D(H) \quad \wedge \\
\nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}'} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor  \quad \wedge \\
\nonumber& & H_{\mathrm{l'}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}'} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
\nonumber& & H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}} \quad \wedge \\
\nonumber& & H_{\mathrm{i}} = {P(H)_{\mathrm{H}}}_{\mathrm{i}} +1 \quad \wedge \\
\nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32 \quad \wedge \\
\nonumber& & H_{\mathrm{f}} = F(H)
\end{eqnarray}
where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
